---
title: "Inference from a mediator"
author: "Macartan"
date: "April 27, 2019"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(gbiqq)
library(dplyr)
```

# A model with possible confounding

We explore a model where a front door type criterion is satisfied: $X\rightarrow M \rightarrow Y$ but $X$ is less likely in cases where $Y=1$ regardless.

With simple $X,Y$ data we greatly underestimate the effect of $X$ on $Y$ because of confounding; with full data on $M$ we do very well; but we also do well even with quite partial data on $M$, eg for 40 out of 20,000 cases.


```{r}
model    <- make_model("X -> M -> Y")
P        <- make_parameter_matrix(model, list(ancestor = c(X="X"), descendent_type = list( Y = "11" )))
model    <- set_parameter_matrix(model, P)
model    <- set_priors(model)
```

Graph it:

```{r}
plot_dag(model)
```



# Draw data

We imagine a true model with a treatment effect of .25 but negative confounding. 


```{r}
# true lambda
true_lambda = c(.8, .2,
                .5, .5,
                .25, .5, 0, .25,
                .25, .5, 0, .25)
           
```

Generate three data frames: full XMY data, XY data only, XY and small M data

```{r}
n <- 20000

data       <- dataXY <- data_partial <- simulate_data(model, n = n, lambda = true_lambda)
dataXY$M   <- NA
data_partial$M[sample(1:n, n - 50)] <- NA


```


Naive analysis:

```{r}
with(data, lm(Y~X)$coef)

```

Note that the the estiamted ATE is too low -- about half of what it should be.




# Calculate posteriors given different types of data


```{r, include = FALSE}

# X, Y data only
updated_XY      <- gbiqq(model, dataXY)

# Full Data
updated_full    <- gbiqq(model, data)

# Partial Data
updated_partial <- gbiqq(model, data_partial)
```


```{r, eval = FALSE}

# X, Y data only
updated_XY       <- gbiqq(model, dataXY)

# Full Data
updated_full     <- gbiqq(model, data)

# Partial Data
updated_partial  <- gbiqq(model, data_partial)
```


# Illustrate joint posteriors 

We'll make a function to graph posteriors:

```{r}



plot_posterior <- function(updated, main = "", prior = FALSE) {
  if(prior) {lambdas <- data.frame(t(replicate(4000, draw_lambda(model)))) 
  } else {
  lambdas <- data.frame(rstan::extract(updated$posterior, pars= "lambdas")$lambdas)
  names(lambdas) <- names(model$lambda_priors)}
  lambdas$ATE      <- with(lambdas, M.M01*(Y.Y01 - Y.Y10) + M.M10*(Y.Y10 - Y.Y01))
  lambdas$confound <- with(lambdas, XY11.X1Y11 - X.X1)
  with(lambdas, plot(confound, ATE, xlim = c(-1,1), ylim = c(-.5, .5), main = main))
  abline(v = -.3, col = "red")
  abline(a = .25, b = 0, col = "red")
  abline(v = mean(lambdas$confound))
  abline(a = mean(lambdas$ATE), b = 0)
}

par(mfrow = c(2,2))

plot_posterior(model, main = "prior", prior = TRUE)
plot_posterior(updated_XY, main = "data on X,Y only")
plot_posterior(updated_full, main = "full data on X, Y and M")
plot_posterior(updated_partial, main = "XY for all, M for some")
```


