---
title: "Posteriors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Posteriors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r setup, message = FALSE, warning = FALSE}
if(!requireNamespace("fabricatr", quietly = TRUE)) {
  install.packages("fabricatr")
}

library(CausalQueries)
library(fabricatr)
library(knitr)
library(ggplot2)
```

# Accessing the posterior

When you update a model using `CausalQueries`, `CausalQueries` generates and updates a `stan` model and saves the posterior distribution over parameters in teh model. 

The basic usage is:

```{r}
data <- data.frame(X = rep(c(0:1), 10), Y = rep(c(0:1), 10))
model <- make_model("X -> Y") |> 
  update_model(data, refresh = 0)
```

The posterior over parameters can be accessed thus:

```{r}
grab(model, "posterior_distribution")
```


# Summary of stan performance

You can access a summary of the parameter values and convergence information as produced by `stan` thus:

```{r}
grab(model, "stan_fit_summary")
```

This summary provides information on the distribution of parameters as well as convergence diagnostics, summarized in the `Rhat` column.

See [stan documentation](https://cran.r-project.org/web/packages/rstan/vignettes/stanfit-objects.html) for details.


# Advanced diagnostics 

If you are interested in advanced diagnostics of performance you can save and access the `raw` stan output. 

```{r}
model <- make_model("X -> Y") |> 
  update_model(data, refresh = 0, keep_fit = TRUE)
```

Note that the summary for this raw output shows the labels used in the generic `stan` model: `lambda` for the vector of parameters, corresponding to the parameters in the parameters dataframe  (`grab(model, "parameters_df")`), and , if saved,  a vector `types` for the causal types (see `grab(model, "causal_types")`) and `w` for the event probabilities (`grab(model, "event_probabilities")`). 

```{r}
model |> grab("stan_fit")
```


You can then  use diagnostic packages such as `bayesplot`.

```{r}
model |> grab("stan_fit") |>
  bayesplot::mcmc_pairs(pars = c("lambdas[3]", "lambdas[4]", "lambdas[5]", "lambdas[6]"))

```
      
```{r}
np <- model |> grab("stan_fit") |> bayesplot::nuts_params()
head(np) |> kable()

model |> grab("stan_fit") |>
  bayesplot::mcmc_trace(pars = "lambdas[5]", np = np) 
```
